Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-12-19 Referee1_Response_fl.tex

\bigskip

\noi  
{\large \bf ``Major'' comments:}
\medskip

\begin{itemize}
\item[\textbf{M1.}] \textit{In the typical simulation setting, the computational effort required to produce the output dominates, and hence one would like to ``make the most'' of the output in terms of analysis. 
Sometimes, a similar motivation can arise in stochastic optimization because there may be a computationally expensive underlying simulation model that must be run to produce the i.i.d.\ realizations that are then used in an SAA approach. 
(While they don't now, the authors could include this as motivation for
their work.)}
\end{itemize}

\noi 
Thank you for this suggestion. 
You are right.
There are cases when generating scenarios is expensive in the stochastic optimization setting as well. 
We changed the introduction based on your, the associate editor, and the other reviewer's comments. 
We specifically added this as a motivation in the first full paragraph on page four of the revised paper. 
\medskip

@@

\begin{itemize}
\item[] \textit{However, perhaps more often, sampling the i.i.d. realizations is computationally cheap compared to solving the resulting SAA instances in stochastic optimization. 
In this latter case, we have a choice between simply increasing the number of batches in a non-overlapping method or using the overlapping batch means approach (with some degree of overlap) in order to reduce variance.
Based on the paper's results, perhaps the computational effort required to achieve a level of variance reduction should guide this choice. 
The authors do point to the fact that warm starts in a stochastic
optimization algorithm can be exploited much more easily when using overlapping batch means than when not using overlapping batches. 
Some attempt to explore this tradeoff would be valuable. 
The computational results use sorting (for the news vendor problem), a dynamic programming algorithm (for a stochastic knapsack model), and a decomposition method (for two-stage stochastic linear programs). 
While I am not suggesting that this exploration needs to be made for all three types of algorithms, even exploring it in the case of the sorting algorithm (where some algorithms are well suited for almost-sorted lists) could provide valuable insight.}
\end{itemize}

\noi 
We studied computational efficiency of the OMRP variance estimators and demonstrated efficiency on the newsvendor problem using insertion sort. 
The results of this study are presented in a new section, \S 5.4 (Computational Efficiency).
We first approached efficiency by the reciprocal of variance multiplied by the computation time, per the associate editor's suggestion. 
Maximizing computational efficiency defined in this manner provided guidelines on selecting the degree of overlap. 
We then looked at how OMRP compared to increasing the number batches in MRP to reach an equivalent variance level.
We examined computational efficiency theoretically---assuming a functional form of the solution time per batch---and also empirically. 
For empirical evaluation, we used the newsvendor problem with insertion sort, based on your suggestions; thank you. 
We found that when effective warm starting is available, overlapping can lead to more efficient estimators than (nonverlapping) MRP, both when the direct nonoverlapping counterpart is used or when the number of batches in MRP is increased to match the variance of the overlapping counterpart.  
We believe the study of computational efficiency significantly improved the paper. 
Thank you.\smallskip

@@

\noindent 
Done. 
\smallskip

\noi 
{\it Note:} This sentence now appears in the second paragraph of \S 3 (Overlapping Multiple Replications Procedure) as a result of restructuring the introduction based on the comments of the associate editor and the other referee. 
\medskip

\begin{itemize}
\item[6.] \textit{p4, l42: Suggest: Partial overlap results in fewer batches, hence fewer optimization \ldots}
\end{itemize}

\noindent 
We revised this paragraph in order to improve exposition. 
We used `smaller number of batches' based on the comments of the other referee in the new sentence. 
\medskip

\noi 
{\it Note:} The revised sentence now appears in the second paragraph of \S 3 (Overlapping Multiple Replications Procedure) as a result of restructuring the introduction based on the comments of the associate editor and the other referee. 
\medskip

@@

\begin{itemize}
\item[12.] \textit{ p14, l23: Please add $j = 1$ to the sum.}
\end{itemize}

\noindent 
Added. Thanks for catching this.
\medskip

\begin{itemize}
\item[13.] \textit{p14, l35: The authors have gone to some length to distinguish different types of $A_j$ terms.
However, this equation has an $\mathbb{E}A_j^2$ term that has been collapsed from a sum, but we don't know ``which $j$''. 
Should there be a $\max_j \mathbb{E}A_j^2$ or some other treatment here?}
\end{itemize}

\noindent 
This is because the center batches are identically distributed. 
As a result, $\mathbb{E}A_j^2$ for all the center batches $j=\lceil \bar{\gamma}^{-1} \rceil, \ldots,n_b - \lceil \bar{\gamma}^{-1} \rceil$ are the same. 
To make this clear, we introduced the notation $A_j^c$ for center batches and $A_j^f$ and $A_j^e$ for the fringe batches that appear at the front and end of each set of nonoverlapping batches, respectively. 
We also emphasized that each front fringe batch has an identically distributed pair at the end fringes, and that all the center batches are identically distributed. 
\medskip

\noi 
{\it Note:} We added brackets to expectations to improve readability as suggested by the other referee.
We also removed the second equality to save some space because this equality's result was also equally expressed in the last inequality. 
\medskip

### 2014-12-19 Referee2_Response_fl.tex

\bigskip

\noi  
{\large \bf General Comments:}
\medskip

\begin{itemize}
\item[1.] \textit{The main contributions of the paper need to be stated more clearly and succinctly.}
\end{itemize}

\noindent 
You are right. 
We changed the introduction to clearly state the main contributions of the paper.  
Please see page four of the revised manuscript. 
We also clarified the motivation for the work, included new contributions based on the study of computational efficiency (contribution (iv)), and provided a comparison with the proceedings version of the paper (based on the editor-in-chief's suggestion).
\medskip

\noi 
As a result of the revised introduction---aimed to clearly state the contributions of the paper---we decided to move the technical discussion on using overlapping batches in simulation output analysis versus assessing solution quality in stochastic optimization to \S 3 (Overlapping Multiple Replications Procedure).
We provided `{\it Note:}'s whenever a sentence was moved when addressing your minor comments below. 
We believe the presentation of the paper significantly improved.  
Thank you.
\medskip

@@

\begin{itemize}
\item[2.] \textit{It is not necessary to go into such detail in the OBM review.}
\end{itemize}

\smallskip

\noindent 
We shortened and streamlined the OBM review, focusing on the main results used in the paper, by combining two previous sections---\S 2.1 (Overlapping Batch Means) and \S 3.1 (Variably Overlapping Batches)---along with other explanations on OBM that appeared later in the paper. 
The combined review is in the background section, \S 2.1 (Batch Means). 
To improve readability, we created two subsections: \S 2.1.1 (Nonoverlapping Batch Means) and \S 2.1.2 (Overlapping Batch Means). 
While we removed a number of equations and shortened the text, we kept some explanations that are necessary for a reader who is not familiar with OBM (e.g., readers with an optimization focus) to follow the main ideas.   
We hope to reach a broader audience this way. 
As a result of these changes: 
\begin{itemize}
	\item[(i)]  instead of former equations (2) and (5), both on overlapping batch means variance estimator, we have only one equation (2);

@@

	\item[(ii)]  the confidence interval estimation using OBM is discussed only once;

	\item[(iii)] the former equation (3) is removed;

	\item[(iv)] the discussion that appeared at the beginning of \S 4.4 (previously: Variance Reduction, now: Asymptotically Lower Variances) is moved to this review section.\medskip

{\it Note:} While making the changes, we realized that notation `$\alpha$' was overloaded. 
It was used in the current equation (3) and the statement of Theorem 2 to represent the relative variances, and it was used as the level of significance in confidence intervals.  
To remedy this issue, we changed `$\alpha$' in equation (3) and Theorem 2 to `$\beta_{\bar{\gamma}}$'. 
\end{itemize}

@@

\noi
Done.
\medskip

\begin{itemize}
\item[] \textit{2, 24: ``under which the resulting gap estimators exhibit reduced variance when compared to their counterparts obtained via nonoverlapping batches.''}
\end{itemize}

\noi
Done.
\medskip

\noi
{\it Note:}  We moved this sentence from the paragraph on page two to later in the introduction to clearly and succinctly state the contributions of the paper. 
It now appears as contribution (ii) on page four of the revised manuscript.
\medskip

@@

\begin{itemize}
\item[] \textit{2, 24 (and elsewhere): Note: I would be careful about the use of the term ``variance reduction'', which usually refers to the class of techniques such as common random numbers, antithetics, importance sampling, etc. 
I view what you're doing in this paper as simply proposing a new estimator whose variance is lower than that of a competing estimator.}
\end{itemize}

\noi
You are right. 
Thanks for pointing this out. 
We corrected the wording throughout the paper, avoiding the use of the term `variance reduction'. 
\medskip

@@

\begin{itemize}
\item[] \textit{2, 27: ``and that the interval \ldots valid. 
We also present''}
\end{itemize}

\noi
We made the change to the first sentence above. 
We reworded the second sentence because it starts a new set of contributions in the revised paper. 
\medskip

\noi
{\it Note:}
The first sentence now appears as part of contribution (ii) and the second sentence as part of contribution (iii) on page four of the revised manuscript. 
\medskip

@@

\begin{itemize}
\item[] \textit{2, 34 (and elsewhere): Might want to boldface all vectors.}
\end{itemize}

\noi
Done.  
\medskip

\begin{itemize}
\item[] \textit{2, 42 (and elsewhere): ``more-restrictive moment'' (hyphenate certain non-adverbial compound adjectives)}
\end{itemize}

\noi
Done.  
We checked for other compound adjectives and changed them according to the Chicago Manual of Style. 
For instance, the manual recommends not to hyphenate `adverb ending in ly + participle or adjective' even it appears before a noun (e.g., maximally overlapping batches); therefore we removed the hyphen from this phrase. 
\medskip

@@

\noi
You are right. 
We removed the comment on the central limit theorem on the minimized sample means as it is not relevant to the paper. 
We also reworded these sentences to improve clarity. 
\medskip

\noi
{\it Note:}
The revised sentences now appear in \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

\begin{itemize}
\item[] \textit{4, 31: What are the desired statistical properties?}
\end{itemize}

\noi
Thanks for pointing out the ambiguity here. 
We reworded this sentence, divided it into two sentences to improve readability, and added further explanations to remove any remaining ambiguities. 
\medskip

@@

\noi
Changed it to ``\dots because we need to solve \ldots '' to improve readability.  
\medskip

\noi
{\it Note:}
This sentence now appears in \S 3 (Overlapping Multiple Replications Procedure). 
\medskip

\begin{itemize}
\item[] \textit{4, 42: ``Partial overlapping results in a smaller number of batches; hence, fewer optimization''}
\end{itemize}

\noi
We revised this section in order to improve exposition. 
We used `smaller number of batches' based on your suggestion in the revised sentence. 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 51: ``discuss overlapping''}
\end{itemize}

\noi
Done.  
\medskip

\begin{itemize}
\item[] \textit{4, 54 (and elsewhere): When you refer to ``variable amounts of overlap'', that is really nothing new. Certainly, any overlapping estimator will have end effects in which the overlap amounts differ from those in the ``middle'' region of the data.}
\end{itemize}

\noi
We meant to say that the degree of overlap is parametrized and this parameter can be chosen by the user. 
We reworded this sentence to reflect this. 
We went through the paper and removed other occurrences of phrases like `variable amounts of overlap', `variably overlapping batches', etc.
We also avoided using these terms in the newly added text. 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 57 (and elsewhere): ``classical case'' refers to nonoverlapping batch means, right?}
\end{itemize}

\noi
No. By `classical case', we meant the use of overlapping batches in simulation output analysis (as opposed to assessment of solution quality in stochastic programming). 
We reworded this sentence. 
%Based on your earlier comments, we avoided the term `variance reduction' in the revised sentence.
We also went through the paper and removed several uses of the word `classical' that referred to nonoverlapping batches.
Any remaining sentences with the word `classical' only refer to the use of overlapping batches in simulation output analysis. 
%In the current version of the paper, `classical case' is used to only indicate  the use of overlapping batches in simulation output analysis. 
\medskip

@@

\begin{itemize}
\item[] \textit{5: You can significantly reduce the discussion on OBM here and in \S 3.1. 
Just let the reader know the main results that you'll use.}
\end{itemize}

\noi
We removed \S 3.1 (Variably Overlapping Batches) from the paper. 
We merged this section with \S 2.1.2 (Overlapping Batch Means), focusing on the results used in the paper. 
We believe the presentation of the paper improved as a result. 
Thank you for this suggestion. 
\medskip

\begin{itemize}
\item[] \textit{6, 13: ``corresponding overlapping batch means''}
\end{itemize}

\noi
Done.  
We removed the word `corresponding' as part of the rewrite but added the word `overlapping' based on your suggestion.
\medskip

@@

\begin{itemize}
\item[] \textit{6, 29: Note that you use ``Note that'' too often.}
\end{itemize}

\noi
We went through the paper and removed most of the ``note that'' phrases. 
We decided to leave some as is (as ``notice that'') because they highlight important characteristics.  
\medskip

@@

\begin{itemize}
\item[] \textit{6, 30: Explicitly say what the d.f.\ in (2) is.}
\end{itemize}

\noi
We changed this sentence and directly referred to the denominator.
We decided to avoid the `degrees of freedom' here. 
This is because we wanted to focus on the degrees of freedom for confidence interval construction, which is an important component of both our theoretical and computational results.\medskip

@@

\noi 
{\it Note:} This section has been restructured based on your comments. 
Equation (2) now gives the general overlapping variance estimator with the parametrized amount of overlap. 
\medskip

\begin{itemize}
\item[] \textit{7, 46: Does $\gamma$ need to be an integer? 
You seem to be using it that way, so maybe state the requirement.}
\end{itemize}

\noi
Yes. We explicitly stated this in a sentence after we introduce $\gamma$.
\medskip

@@

\begin{itemize}
\item[] \textit{12, 52: ``$n_l VD_l$ has been studied''}
\end{itemize}

\noi
Done.
\medskip

\begin{itemize}
\item[] \textit{13, 12 (and elsewhere): I may have missed this, but has the notation $\sigma_{\hat{x}}$ been established yet?}
\end{itemize}

\noi
This notation was defined before Theorem 1. 
We moved the definitions of $\sigma_{\hat{x}}$ and $\mu_{\hat{x}}$ to the end of \S 4.2 (Nonoptimized Counterparts) to improve clarity and added explanations to provide context. 
\medskip

@@

\begin{itemize}
\item[] \textit{In any case, some of the verbiage in this paragraph is a bit repetitive.}
\end{itemize}

\noi
We moved the discussion in the first half of this paragraph to the new, combined OBM review in \S 2.1.2 (Overlapping Batch Means). 
We referred to equation (3) here to remind the readers the basis of the proof.
The second half of this paragraph deals with the batching structure, and this discussion is moved to the paragraph after the theorem's proof. 
\medskip

@@

\begin{itemize}
\item[] \textit{13, 53: The reader might need another step besides C-S to get that inequality.}
\end{itemize}

\noi
This is a direct application of the Cauchy-Schwartz inequality $\left(\sum_{j=1}^{n}x_j y_j\right)^2 \leq \sum_{j=1}^{n}x_j^2 \sum_{j=1}^{n}y_j^2$, where the $y_j$ are all set to $1$. 
We added a brief explanation after this inequality to make it clear. 
\medskip

\begin{itemize}
\item[] \textit{15, 25: Do you discuss $t_{d_l,\alpha}$ at all in Theorem 3?}
\end{itemize}

\noi
No, we had not.   
We added explanations in the proof Theorem 3, discussing how $t_{d_l(k_l - 1),\alpha}$ factors in the proof. 
\medskip

\noi 
{\it Note:}
While making this revision we noticed a typo. 
We changed $t_{d_l,\alpha}$ to $t_{d_l(k_l - 1),\alpha}$, where $k_l=\left\lfloor \frac{n_l}{m_l} \right\rfloor$ is the number of batches used in the nonoverlapping counterpart. 
\medskip

\newpage
\begin{itemize}
\item[] \textit{18: Your figures are very difficult to read, even with the pdf file.}
\end{itemize}

### 2014-12-19 omrp_lb_fl.tex
CHANGED THE FILES
For the first look revision...

\begin{abstract}
\noindent Overlapping Batch Means (OBM) has long been used in simulation as a method of reusing data to generate variance estimators with asymptotically lower variance.
In this paper, we apply the OBM method to stochastic programming by formulating a variant of the multiple replications procedure used for assessing solution quality.
We give conditions under which the resulting optimality gap point estimators are strongly consistent, interval estimators are asymptotically valid, and the variance estimators have the same asymptotically lower variances relative to their nonoverlapping counterparts as the classical OBM estimators \citep{Meketon1984,Welch1987}.
Computational experiments for several test problems are presented, examining the small-sample behavior and the computational efficiency of the overlapping batches method in this context.\medskip

@@

Unless $f$ has a simple structure or the number of realizations of $\xit$ is small, (\ref{eq:sto_prog}) typically cannot be solved exactly. 
An approximation of (\ref{eq:sto_prog}) can be generated by sampling from the distribution of $\xit$.
While other sampling schemes are possible, let $\xit^1, \xit^2, \dots, \xit^m$ be an independent and identically distributed (i.i.d.) sample from the distribution of $\xit$.
A sampling approximation of (\ref{eq:sto_prog}), often called the sample average approximation, is given by
\begin{align} \tag{SP$_m$} \label{eq:sto_prog_m}
	z_m^* & = \min_{\x \in X} \frac{1}{m} \sum_{i=1}^m f(\x,\xiti).
\end{align}
We denote an optimal solution to (\ref{eq:sto_prog}) as $\xs$ and an optimal solution to (\ref{eq:sto_prog_m}) as $\xs_m$.

@@

Overlapping may increase the computational burden in the stochastic optimization setting.
By jointly considering computation time and variance, we have examined the computational efficiency of OMRP, and we have empirically demonstrated  efficiency by solving a newsvendor problem using insertion sort.
Our analysis illustrates that when warm starting is effective, computational efficiency gains can be observed across a wide range of overlapping degrees.
Even when warm starting is less effective, efficiency slightly increases for low degrees of overlap. 
Maximizing efficiency and then minimizing total computation time suggests selecting the lowest degree of overlap across all the degrees with the highest efficiency for a given problem and solution method (with warm starting) combination.

@@

Overall, the results in this paper suggest that OMRP could be an efficient method to estimate optimality gaps for a class of stochastic programs when warm starting is available and/or when generating the observations is costly.

>>>>

\begin{abstract}
\noindent Overlapping Batch Means (OBM) has long been used in simulation as a method of reusing data to generate variance estimators with asymptotically lower variance.
In this paper, we apply the OBM method to stochastic programming by formulating a variant of the multiple replications procedure used for assessing solution quality.
We give conditions under which the resulting optimality gap point estimators are strongly consistent, interval estimators are asymptotically valid, and the variance estimators have the same asymptotically lower variances relative to their nonoverlapping counterparts as the classical OBM estimators \citep{Meketon1984,Welch1987}.
We investigate computational efficiency, a combined measure of variance and computation time, providing guidelines on the degree of overlap.  
Numerical experiments on several test problems are presented, examining the small-sample behavior and the empirical computational efficiency of the overlapping batches method in this context.  
\medskip

@@

Unless $f$ has a simple structure or the number of realizations of $\xit$ is small, (\ref{eq:sto_prog}) typically cannot be solved exactly. 
An approximation of (\ref{eq:sto_prog}) can be generated by sampling from the distribution of $\xit$.
While other sampling schemes are possible, let $\xit^1, \xit^2, \dots, \xit^m$ be an independent and identically distributed (i.i.d.) sample from the distribution of $\xit$.
A sampling approximation of (\ref{eq:sto_prog}), often called the sample average approximation, is given by\vspace*{-0.07in}
\begin{align} \tag{SP$_m$} \label{eq:sto_prog_m}
	z_m^* & = \min_{\x \in X} \frac{1}{m} \sum_{i=1}^m f(\x,\xiti).
\end{align}
We denote an optimal solution to (\ref{eq:sto_prog}) as $\xs$ and an optimal solution to (\ref{eq:sto_prog_m}) as $\xs_m$.

@@

Overlapping may increase the computational burden in the stochastic optimization setting.
By jointly considering computation time and variance, we have examined the computational efficiency of OMRP, and we have empirically demonstrated  efficiency by solving a newsvendor problem using insertion sort.
Our analysis illustrates that when warm starting is effective, computational efficiency gains can be observed across a wide range of overlapping degrees.
Even when warm starting is less effective, efficiency slightly increases for low degrees of overlap. 
Maximizing efficiency and then minimizing total computation time, we recommend  selecting the lowest degree of overlap across all the degrees with the highest efficiency for a given problem and solution method (with warm starting) combination.

@@

Overall, the results in this paper suggest that OMRP could be an efficient method to estimate optimality gaps for a class of stochastic programs when warm starting is available and/or generating the observations is costly.
